Probabilities of polar alignment in circumbinary disks

### 2024-12-19 src/tex/ms.tex
Add RKF references

%\section{Runge-Kutta-Fehlburg method}
\section{Secular Approximation}
\label{sec:rk}
\citet[Equations (7-10)]{martin2019} also provide a set of coupled first-order differential equations that describe the evolution of $e_{\rm b}$ along with the direction of the disk angular momentum $(\ell_x, \ell_y, \ell_z)$, based on previous work by \citet{farago2010}. We solve these equations using the Runge-Kutta-Fehlberg (RKF) method \TJ{ref?} -- a numerical method similar to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size. This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period \citep[e.g.][]{naoz2016}. However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than running $n$-body simulations (see Section \ref{sec:reb}).

@@

Figure \ref{fig:rkf} shows that, while the RKF and {\sc rebound} simulations produce similar results, they are not identical. Specifically noticeable are points sampled by the MC algorithm that lie outside the RKF libration region (with a higher inclination), but are found by {\sc rebound} to librate. There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce prograde precession -- but they are fewer and not as noticeable by eye. This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary \TJ{(is this the correct way to word this?)}, while in this section we are treating the full $n$-body dynamics. The two schemes converge when $a \gg a_\text{b}$, and we can eliminate this discrepancy by increasing $a$ (e.g., $a=25a_\text{b}$ compared to our nominal choice of $5a_\text{b}$). In general, the choice of small $a/a_\text{b}$ changes the shape of the parameter space that allows polar orbits when compared to those found in Section \ref{sec:rk}, but not enough to have a significant effect on $f_\text{polar}$. In fact, this bias is small enough that -- given uncertainties introduced by our limited number of samples -- we cannot distinguish between the polar fractions inferred from the two methods. Because the secular approximation used in Sections \ref{sec:analytic} \& \ref{sec:rk} require significantly lower computational cost, and the differences in results are insignificant, we choose to focus on the secular approximation results in the following sections.

>>>>

%\section{Runge-Kutta-Fehlburg method}
\section{Secular Approximation}
\label{sec:rk}
\citet[Equations (7-10)]{martin2019} also provide a set of coupled first-order differential equations that describe the evolution of $e_{\rm b}$ along with the direction of the disk angular momentum $(\ell_x, \ell_y, \ell_z)$, based on previous work by \citet{farago2010}. We solve these equations using the Runge-Kutta-Fehlberg (RKF, \citealt{fehlberg1969}, see \citealt{hairer2000} p. 177 for implementation) -- a numerical method similar to the common 4th order Runge-Kutta integrator, but that computes a 5th order solution to allow for a variable step size. This method treats the binary in the quadrupole approximation, meaning that it is not sensitive to the effects of the binary mass fraction or orbital dynamics that occur on timescales less than the binary orbital period \citep[e.g.][]{naoz2016}. However, the computational cost is extremely low; we are able to run $10^4$-$10^5$ simulations per second -- about 10,000 times faster than running $n$-body simulations (see Section \ref{sec:reb}).

@@

Figure \ref{fig:rkf} shows that, while the RKF and {\sc rebound} simulations produce similar results, they are not identical. Specifically noticeable are points sampled by the MC algorithm that lie outside the RKF libration region (with a higher inclination), but are found by {\sc rebound} to librate. There are also points with the opposite discrepancy -- those that lie in the RKF libration region but produce prograde precession -- but they are fewer and not as noticeable by eye. This discrepancy arises because the secular approximation averages the central potential over the orbital period of the binary, while in this section we are treating the full $n$-body dynamics. The two schemes converge when $a \gg a_\text{b}$, and we can eliminate this discrepancy by increasing $a$ (e.g., $a=25a_\text{b}$ compared to our nominal choice of $5a_\text{b}$). In general, the choice of small $a/a_\text{b}$ changes the shape of the parameter space that allows polar orbits when compared to those found in Section \ref{sec:rk}, but not enough to have a significant effect on $f_\text{polar}$. In fact, this bias is small enough that -- given uncertainties introduced by our limited number of samples -- we cannot distinguish between the polar fractions inferred from the two methods. Because the secular approximation used in Sections \ref{sec:analytic} \& \ref{sec:rk} require significantly lower computational cost, and the differences in results are insignificant, we choose to focus on the secular approximation results in the following sections.
